---
title: "Mobility"
author: "Jeremy Leipzig"
date: "April 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(smacof)
library(MASS)
library(knitr)
```

Career mobility data represented as profession matrix from 1962 (Y-axis) to 1973 (X-axis). Multidimensional scaling offers some interesting, if not a complete view of these changes.

Some careers offer more mobility overall than others, and these should be represented by centrality in an MDS plot. Careers which are terminal should reside on the periphery.

Certain career tracks form clusters of related professions. 

```{r matrixsetup, message=FALSE}
tb<-read.table("MOBILITY.txt")
as.matrix(tb[,-1])->mob_mat
pronames<-c("pro_se","pro_sal","mgr","sales_oth","props","clerks","sales_retail","crafts_mfr","crafts_oth","crafts_cnstr","srvc","oper_oth","oper_mfr","labor_mfr","labor_oth","farmers","farm_labor")
dimnames(mob_mat)<-list(pronames,pronames)
knitr::kable(mob_mat)

```

## Using dist and isoMDS
This procedure attempts to force asymmterical distances into symmetrical ones. We will treat these as non-metric distances.

```{r discmd, message=FALSE}
diag(mob_mat)<-0
d <- dist(1-sqrt(mob_mat))
fit <- isoMDS(d, k=2)
fit
points<-as.data.frame(fit$points)
points$profession<-row.names(points)
ggplot(points,aes(V1,V2,label=profession))+geom_point()+geom_text(aes(x= V1+2, y = V2 + 1))+xlab("X1")+ylab("X2")
```

## Using driftVectors
The driftVectors package attempts to create a symmetrical profile compatible with an MDS plot while storing vectors that relate asymmetry.

```{r drift, message=FALSE}
mob_scale<-scale(mob_mat)-min(scale(mob_mat))
mob_scale_dist<-2.5-sqrt(mob_scale)
driftVectors(mob_scale_dist,type="ratio")
plot(driftVectors(mob_scale_dist,type="ratio"),xlim =c(-1,1), ylim=c(-1,1), col.drift = "blue")
```

## Conclusions

* The labor professions are more closely tied to each other than their respective economic units

* Self-employment and farming are destination professions
